\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
This collection of articles was a result of a long journey that started in 2015 when I joined the PhD program at CLASP. Producing such a piece is not a one-person job. I was fortunate to have the support of my supervisors, colleagues, friends and family along the way.

First of all, I would like to recognize the invaluable assistance of my supervisors. 
I want to express my most profound appreciation to Simon Dobnik, my primary advisor, for his steady support of my trembling steps on this path. A large body of this work became possible with his direct guidance. 
I wish to acknowledge the efforts of Staffan Larsson, my second advisor, whose feedback helped me at various times during this project. 
I am incredibly grateful to John Kelleher, my third advisor, our long passionate conversations shaped many ideas presented here.
In addition to the help I received from supervisors, I was lucky for being able to seek advice from other fantastic people. 
I am indebted to Richard Johansson for his counselling which led me to apply for the PhD program. I had the privilege to receive his support since I was a master student. 
I am also grateful to Stergios Chatzikyriakidis for his encouraging feedback. I have always been able to rely on his friendly advice.
Also, I want to thank Lena Eriksson for her counselling. Her thoughtful advice proved to have a significant influence on the success of my studies.
I wish to thank, Desmond Elliott, my thesis assessor, for his constructive criticism and helpful feedback on an earlier version of the thesis.

I want to extend my sincere thanks to friends and colleagues for building such an excellent social and research environment. 
Shalom Lappin, for his unparalleled support and care;
Robin Cooper, for sharing extensive knowledge in our bi-weekly meetings at the language and perception reading group; 
Jean-Philippe Bernardy, for his constructive advice from time to time;
Yuri Bizzoni, for being a colleague and a friend with a significant positive atmosphere around him. Also, thanks to Johan Gross and Julia Forsberg, with Yuri, we had a relaxed PhD life at the office! 
Same regards to Adam Ek and Vidya Somashekarappa my officemates for almost a year. Although we only shared office for a short period, we made unforgettable memories. I literally received a collection of life lessons from Vidya and Adam after their thoughtful prank! 
I also wish to thank all the incredible people who made Dicksonsgatan 4 like a family house for me. Haris Themistocleous, Chris Howes, Ellen Breitholtz, Sharid Loáiciga, Asad Sayeed, Sandro Maskharashvili, Rasmus Blanck, Wafia Adouane, Vlad Maraev, Bill Noble, Sylvie Saget, Kathrein Abu Kwaik, Susanna Myyry, Iines Turunen, Aram Karimi. We had a lot of meaningful spontaneous conversations in the kitchen, corridors and the meeting room. I learned something memorable from each of you. I won't miss a chance if I get to work with you again. 

The extent of the fantastic colleagues that I wish to thank them for their company goes beyond the circle of CLASP, at GU and Chalmers. Conversation with these friendly faces kept me on the track! I want to thank Mio Cibulka, Leila El-Alti, Richard Endörfer, David Alfter, Luis Nieto Piña, Ildikó Pilán, Prasanth Kolachina, Herbert Lange.  

I cannot begin to express my thanks to my family and friends whose support cannot be overestimated. These extraordinary people kept me sane during my PhD life. In Sweden, I had incredible times with Alireza Pazirandeh, Ala Pazirandeh, Niklas Arvidsson and My Högblom. 
I wish to thank my sisters, Marzieh and Mona, and my mother and father, Shokoufeh and Siamack, for profound belief in my work and my abilities. 

Lastly, I would like to thank my wife, Boshra Khoshnevis (without forgetting the help of Biloo and Ghoughies). She is the reason that kept me going on. The completion of my dissertation without her support would not be possible.



